### 2023-11-15 production/prod_prep-ingresos.Rmd

---
title: "prod_prep_ingresos"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
      number_sections: yes
    code_folding: hide
    number_sections: yes
editor_options:
  chunk_output_type: console
---

# Setup

```{r setup}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
options(scipen=9999) # desactivar notacion cientifica
remove(list = ls()) #limpieza del entorno de trabajo
```

```{css, echo=FALSE}
# /* expandir el espacio del documento*/
body .main-container {
      max-width: 1600px !important;
      width: 1600px !important;
    }
#/* expandir width de la TOC*/
div.tocify {
    width: 22% !important;
    max-width: 331px !important;
    max-height: 102% !important;
}
```

# Download data
```{r download data, eval=FALSE, include=FALSE}
#download and save data

file.exists(here::here("input/data-original/elsoc_wide.Rdata"))

load(url("https://dataverse.harvard.edu/api/access/datafile/7245115")) #ELSOC
save(elsoc_wide_2016_2022.2,file = here::here("input/data-original/elsoc_wide.Rdata"))
```

## Individual data

```{r}
remove(list = ls()) # clean workspace
# .rs.restartR() #restart R session
if (!require("pacman")) install.packages("pacman") # install pacman
# load libraries
pacman::p_load(dplyr,readxl,sjmisc, sjlabelled, questionr, car,here, summarytools)
load(here::here("input/data-original/elsoc_wide.Rdata"))

elsoc_wide_2016_2022.2[elsoc_wide_2016_2022.2 ==-999] <- NA
elsoc_wide_2016_2022.2[elsoc_wide_2016_2022.2 ==-888] <- NA
elsoc_wide_2016_2022.2[elsoc_wide_2016_2022.2 ==-777] <- NA
elsoc_wide_2016_2022.2[elsoc_wide_2016_2022.2 ==-666] <- NA

# ingresos__________________________________________________
ingresos <- 
elsoc_wide_2016_2022.2 %>% 
  #create new variables
  select(m30_w01,
         m30_w02,
         m30_w03,
         m30_w04,
         m30_w06,
         m30b_w05,
         m30b_w06,
         m29_w01,
         m29_w02,
         m29_w03,
         m29_w04,
         m29_w05,
         m29_w06,
         ) %>% 
  sjlabelled::drop_labels(., drop.na = FALSE)
```

```{r echo=FALSE}
print(dfSummary(ingresos, headings = FALSE, graph.col = FALSE), method = "render")
```

```{r}
# Household income_________________________________________

#Impute midpoint of income ranges
elsoc_wide_2016_2022.2$m30_rec_w01 <-
as.numeric(car::recode(elsoc_wide_2016_2022.2$m30_w01,
           "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=2700000;NA=NA;c(-888,-999)=NA"))

elsoc_wide_2016_2022.2$m30_rec_w02 <-
as.numeric(car::recode(elsoc_wide_2016_2022.2$m30_w02,
           "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=2700000;NA=NA;c(-888,-999)=NA"))

elsoc_wide_2016_2022.2$m30_rec_w03 <-
as.numeric(car::recode(elsoc_wide_2016_2022.2$m30_w03,
           "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=2700000;NA=NA;c(-888,-999)=NA"))

elsoc_wide_2016_2022.2$m30_rec_w04 <-
as.numeric(car::recode(elsoc_wide_2016_2022.2$m30_w04,
           "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=2700000;NA=NA;c(-888,-999)=NA"))

elsoc_wide_2016_2022.2$m30_rec_w06 <-
as.numeric(car::recode(elsoc_wide_2016_2022.2$m30_w06,
           "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=2700000;NA=NA;c(-888,-999)=NA"))

#Impute midpoint of income ranges (2021)
elsoc_wide_2016_2022.2$m30b_rec_w05 <-
as.numeric(car::recode(elsoc_wide_2016_2022.2$m30b_w05,
           "1=125000;2=300000;3=400000;4=575000;5=70000;NA=NA;c(-888,-999)=NA"))

elsoc_wide_2016_2022.2$m30b_rec_w06 <-
as.numeric(car::recode(elsoc_wide_2016_2022.2$m30b_w06,
           "1=125000;2=300000;3=400000;4=575000;5=70000;NA=NA;c(-888,-999)=NA"))

sjmisc::frq(elsoc_wide_2016_2022.2$m30b_rec_w05)

#Recode DK/DA of Income to NA
elsoc_wide_2016_2022.2$m29_rec_w01 <-
  as.numeric(car::recode(elsoc_wide_2016_2022.2$m29_w01,"c(-888,-999)=NA"))

elsoc_wide_2016_2022.2$m29_rec_w02 <-
  as.numeric(car::recode(elsoc_wide_2016_2022.2$m29_w02,"c(-888,-999)=NA"))

elsoc_wide_2016_2022.2$m29_rec_w03 <-
  as.numeric(car::recode(elsoc_wide_2016_2022.2$m29_w03,"c(-888,-999)=NA"))

elsoc_wide_2016_2022.2$m29_rec_w04 <-
  as.numeric(car::recode(elsoc_wide_2016_2022.2$m29_w04,"c(-888,-999)=NA"))

elsoc_wide_2016_2022.2$m29_rec_w05 <-
  as.numeric(car::recode(elsoc_wide_2016_2022.2$m29_w05,"c(-888,-999)=NA"))

elsoc_wide_2016_2022.2$m29_rec_w06 <-
  as.numeric(car::recode(elsoc_wide_2016_2022.2$m29_w06,"c(-888,-999)=NA"))

#replace NA of income with new imputed variable
elsoc_wide_2016_2022.2$m29_imp_w01 <- 
  ifelse(test = !is.na(elsoc_wide_2016_2022.2$m29_rec_w01),
         yes =  elsoc_wide_2016_2022.2$m29_rec_w01,
         no =  elsoc_wide_2016_2022.2$m30_rec_w01)
summary(elsoc_wide_2016_2022.2$m29_imp_w01)

elsoc_wide_2016_2022.2$m29_imp_w02 <- 
  ifelse(test = !is.na(elsoc_wide_2016_2022.2$m29_rec_w02),
         yes =  elsoc_wide_2016_2022.2$m29_rec_w02,
         no =  elsoc_wide_2016_2022.2$m30_rec_w02)
summary(elsoc_wide_2016_2022.2$m29_imp_w02)

elsoc_wide_2016_2022.2$m29_imp_w03 <- 
  ifelse(test = !is.na(elsoc_wide_2016_2022.2$m29_rec_w03),
         yes =  elsoc_wide_2016_2022.2$m29_rec_w03,
         no =  elsoc_wide_2016_2022.2$m30_rec_w03)
summary(elsoc_wide_2016_2022.2$m29_imp_w03)

elsoc_wide_2016_2022.2$m29_imp_w04 <- 
  ifelse(test = !is.na(elsoc_wide_2016_2022.2$m29_rec_w04),
         yes =  elsoc_wide_2016_2022.2$m29_rec_w04,
         no =  elsoc_wide_2016_2022.2$m30_rec_w04)
summary(elsoc_wide_2016_2022.2$m29_imp_w04)

elsoc_wide_2016_2022.2$m29_imp_w05 <- 
  ifelse(test = !is.na(elsoc_wide_2016_2022.2$m29_rec_w05),
         yes =  elsoc_wide_2016_2022.2$m29_rec_w05,
         no =  elsoc_wide_2016_2022.2$m30b_rec_w05)
summary(elsoc_wide_2016_2022.2$m29_imp_w05)

elsoc_wide_2016_2022.2$m29_imp_w06 <- 
  ifelse(test = !is.na(elsoc_wide_2016_2022.2$m29_rec_w06),
         yes =  elsoc_wide_2016_2022.2$m29_rec_w06,
         no =  elsoc_wide_2016_2022.2$m30_rec_w06)
summary(elsoc_wide_2016_2022.2$m29_imp_w06)

###########

elsoc_wide_2016_2022.2$m29_imp_w05 <- 
  ifelse(test = is.na(elsoc_wide_2016_2022.2$m29_imp_w05),
         yes =  elsoc_wide_2016_2022.2$m30b_rec_w05,
         no =  elsoc_wide_2016_2022.2$m29_imp_w05)
summary(elsoc_wide_2016_2022.2$m29_imp_w05)

elsoc_wide_2016_2022.2$m29_imp_w06 <- 
  ifelse(test = is.na(elsoc_wide_2016_2022.2$m29_imp_w06),
         yes =  elsoc_wide_2016_2022.2$m30b_rec_w06,
         no =  elsoc_wide_2016_2022.2$m29_imp_w06)
summary(elsoc_wide_2016_2022.2$m29_imp_w06)

# deflate at each year's prices
#elsoc_wide_2016_2022.2$deflactor <-
#  with(elsoc_wide_2016_2022.2, case_when(
#    ola == 2016 ~ 113.88 / 123.82,
#    ola == 2017 ~ 116.46 / 123.82,
#    ola == 2018 ~ 119.45 / 123.82,
#    ola == 2019 ~ 123.82 / 123.82
#  ))

# Per capita household income:
elsoc_wide_2016_2022.2$ing_pc_w01 <- 
  (elsoc_wide_2016_2022.2$m29_imp_w01/elsoc_wide_2016_2022.2$nhogar1_w01)

elsoc_wide_2016_2022.2$ing_pc_w02 <- 
  (elsoc_wide_2016_2022.2$m29_imp_w02/elsoc_wide_2016_2022.2$m46_nhogar_w02)

elsoc_wide_2016_2022.2$ing_pc_w03 <- 
  (elsoc_wide_2016_2022.2$m29_imp_w03/elsoc_wide_2016_2022.2$m54_w03)

elsoc_wide_2016_2022.2$ing_pc_w04 <- 
  (elsoc_wide_2016_2022.2$m29_imp_w04/elsoc_wide_2016_2022.2$m54_w04)

elsoc_wide_2016_2022.2$ing_pc_w05 <- 
  (elsoc_wide_2016_2022.2$m29_imp_w05/elsoc_wide_2016_2022.2$m54_w05)

elsoc_wide_2016_2022.2$ing_pc_w06 <- 
  (elsoc_wide_2016_2022.2$m29_imp_w06/elsoc_wide_2016_2022.2$m54_w05)

elsoc_wide_2016_2022.2$ing_pc_w01 <-
sjlabelled::set_label(x = elsoc_wide_2016_2022.2$ing_pc_w01,
                      label = "Household income per capita 2016")

sjmisc::descr(elsoc_wide_2016_2022.2$ing_pc_w01)

# Compute income quintiles
elsoc_wide_2016_2022.2 <- elsoc_wide_2016_2022.2 %>% 
  mutate(quintil_w01 = ntile(-desc(ing_pc_w01), 5))

elsoc_wide_2016_2022.2 <- elsoc_wide_2016_2022.2 %>% 
  mutate(quintil_w02 = ntile(-desc(ing_pc_w02), 5))

elsoc_wide_2016_2022.2 <- elsoc_wide_2016_2022.2 %>% 
  mutate(quintil_w03 = ntile(-desc(ing_pc_w03), 5))

elsoc_wide_2016_2022.2 <- elsoc_wide_2016_2022.2 %>% 
  mutate(quintil_w04 = ntile(-desc(ing_pc_w04), 5))

elsoc_wide_2016_2022.2 <- elsoc_wide_2016_2022.2 %>% 
  mutate(quintil_w05 = ntile(-desc(ing_pc_w05), 5))

elsoc_wide_2016_2022.2 <- elsoc_wide_2016_2022.2 %>% 
  mutate(quintil_w06 = ntile(-desc(ing_pc_w06), 5))

elsoc_wide_2016_2022.2$quintil_w01 <- 
  factor(elsoc_wide_2016_2022.2$quintil_w01,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

elsoc_wide_2016_2022.2$quintil_w02 <- 
  factor(elsoc_wide_2016_2022.2$quintil_w02,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

elsoc_wide_2016_2022.2$quintil_w03 <- 
  factor(elsoc_wide_2016_2022.2$quintil_w03,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

elsoc_wide_2016_2022.2$quintil_w04 <- 
  factor(elsoc_wide_2016_2022.2$quintil_w04,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

elsoc_wide_2016_2022.2$quintil_w05 <- 
  factor(elsoc_wide_2016_2022.2$quintil_w05,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

elsoc_wide_2016_2022.2$quintil_w06 <- 
  factor(elsoc_wide_2016_2022.2$quintil_w06,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

#reverse quintile, reference level is the highest quintile
#elsoc_wide_2016_2022.2$quintil <- forcats::fct_rev(elsoc_wide_2016_2022.2$quintil)

elsoc_wide_2016_2022.2$quintil_w01 <- 
sjlabelled::set_label(x = elsoc_wide_2016_2022.2$quintil_w01,
                      label = "Household income quintile per capita")

sjmisc::frq(elsoc_wide_2016_2022.2$quintil_w01)

#include new quintile category with missing cases
elsoc_wide_2016_2022.2$quintil1_w01<-
  car::recode(elsoc_wide_2016_2022.2$quintil_w01, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

#include new quintile category with missing cases
elsoc_wide_2016_2022.2$quintil1_w02<-
  car::recode(elsoc_wide_2016_2022.2$quintil_w02, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

#include new quintile category with missing cases
elsoc_wide_2016_2022.2$quintil1_w03<-
  car::recode(elsoc_wide_2016_2022.2$quintil_w03, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

#include new quintile category with missing cases
elsoc_wide_2016_2022.2$quintil1_w04<-
  car::recode(elsoc_wide_2016_2022.2$quintil_w04, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

#include new quintile category with missing cases
elsoc_wide_2016_2022.2$quintil1_w05<-
  car::recode(elsoc_wide_2016_2022.2$quintil_w05, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

#include new quintile category with missing cases
elsoc_wide_2016_2022.2$quintil1_w06<-
  car::recode(elsoc_wide_2016_2022.2$quintil_w06, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

#elsoc_wide_2016_2022.2$quintil1 <- factor(elsoc_wide_2016_2022.2$quintil1, c("Q1","Q2","Q3","Q4","Q5","QNA"))

elsoc_wide_2016_2022.2$quintil1_w01 <- 
sjlabelled::set_label(x = elsoc_wide_2016_2022.2$quintil1_w01,
                      label = "Household income quintile per capita (NA)") 
sjmisc::frq(elsoc_wide_2016_2022.2$quintil1_w01)

quintiles <- elsoc_wide_2016_2022.2 %>% select(quintil_w01, quintil_w02, quintil_w03, quintil_w04, quintil_w05, quintil_w06)
quintiles$na_count <- rowSums(is.na(quintiles))
round(prop.table(table(quintiles$na_count))*100,2)
```

```{r echo=FALSE}
print(dfSummary(quintiles, headings = FALSE, graph.col = FALSE), method = "render")
```