### 2023-11-15 production/prod_prep.Rmd

---
title: "Preparación de datos"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
      number_sections: yes
    code_folding: hide
    number_sections: yes
editor_options:
  chunk_output_type: console
---

# Setup

```{r setup}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
options(scipen=9999) # desactivar notacion cientifica
remove(list = ls()) #limpieza del entorno de trabajo
```

```{css, echo=FALSE}
# /* expandir el espacio del documento*/
body .main-container {
      max-width: 1600px !important;
      width: 1600px !important;
    }
#/* expandir width de la TOC*/
div.tocify {
    width: 22% !important;
    max-width: 331px !important;
    max-height: 102% !important;
}
```

# Download data
```{r download data, eval=FALSE, include=FALSE}
#download and save data

file.exists(here::here("input/data-original/elsoc_long.Rdata"))

load(url("https://dataverse.harvard.edu/api/access/datafile/7245118")) #ELSOC
save(elsoc_long_2016_2022.2,file = here::here("input/data-original/elsoc_long.Rdata"))

cit17<- read.table(url("https://dataverse.harvard.edu/api/access/datafile/5216836"), 
                   header = T, sep = "\t", fill = TRUE)
save(cit17,file = here::here("input/data-original/cit17.Rdata"))
cit18<- read.table(url("https://dataverse.harvard.edu/api/access/datafile/5216941"), 
                   header = T, sep = "\t", fill = TRUE)
save(cit18,file = here::here("input/data-original/cit18.Rdata"))
cit19<- read.table(url("https://dataverse.harvard.edu/api/access/datafile/5216990"), 
                   header = T, sep = "\t", fill = TRUE)
save(cit19,file = here::here("input/data-original/cit19.Rdata"))

# Chilean CENSUS data (2017)____________________________________________________
 options(timeout=100*100)
download.file(url = "https://github.com/jciturras/data-chile/raw/main/data/censo-2017/censo-2017.RData",
              destfile = here::here("input/data-original/censo17.Rdata"))

# Chilean Socioeconomic Household Survey (CASEN)_______________________________

# Year 2017_______________________________________________
download.file(url = "https://github.com/jciturras/data-chile/raw/main/data/casen-2017/casen2017.RData",
              destfile = here::here("input/data-original/casen17.Rdata"))
# Year 2020_______________________________________________
# Download the original data from:
# http://observatorio.ministeriodesarrollosocial.gob.cl/storage/docs/casen/2020/Casen_en_Pandemia_2020_STATA.dta.zip 
# note: we changed the original name of the file -> casen20.dta 
casen20 <- 
  haven::read_dta(file = here::here("input/data-original/casen20.dta"))
# save the original data in R format 
save(casen20,file = here::here("input/data-original/casen20.Rdata"))
```

## Individual data

```{r}
remove(list = ls()) # clean workspace
# .rs.restartR() #restart R session
if (!require("pacman")) install.packages("pacman") # install pacman
# load libraries
pacman::p_load(dplyr,readxl,sjmisc, sjlabelled, questionr, car,here)
load(here::here("input/data-original/elsoc_long.Rdata"))

elsoc_long_2016_2022.2[elsoc_long_2016_2022.2 ==-999] <- NA
elsoc_long_2016_2022.2[elsoc_long_2016_2022.2 ==-888] <- NA
elsoc_long_2016_2022.2[elsoc_long_2016_2022.2 ==-777] <- NA
elsoc_long_2016_2022.2[elsoc_long_2016_2022.2 ==-666] <- NA

# Attitudes toward inmigrants__________________________________________________
elsoc_long_2016_2022.2 <- 
elsoc_long_2016_2022.2 %>% 
  #create new variables
  mutate(just_educ=d02_02,
         just_pension=d02_01,
         just_salud=d02_03,
         des_perc=c18_11,
         sexo=m0_sexo,
         merit_effort=c18_09,
         merit_talent=c18_10,
         ahead_family=r12_02,
         ahead_educ=r12_03,
         ahead_ambition=r12_04,
         ahead_hardwork=r12_05) %>% 
  sjlabelled::drop_labels(., drop.na = FALSE)

#
sjmisc::frq(elsoc_long_2016_2022.2$just_educ)
elsoc_long_2016_2022.2$just_educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$just_educ, 
                      label = "Education distributive justice")

#
elsoc_long_2016_2022.2$just_pension <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$just_pension, 
                      label = "Pension distributive justice")

#
elsoc_long_2016_2022.2$just_salud <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$just_salud, 
                      label = "Health distributive justice")

#
elsoc_long_2016_2022.2$des_perc <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$des_perc, 
                      label = "Income differences are too large")

#
sjmisc::frq(elsoc_long_2016_2022.2$merit_effort)
elsoc_long_2016_2022.2$merit_effort <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$merit_effort, 
                      label = "People are rewarded for their efforts")

#
sjmisc::frq(elsoc_long_2016_2022.2$merit_talent)
elsoc_long_2016_2022.2$merit_talent <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$merit_talent, 
                      label = "People are rewarded for their intelligence")

#
sjmisc::frq(elsoc_long_2016_2022.2$ahead_family)
elsoc_long_2016_2022.2$ahead_family <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ahead_family, 
                      label = "Coming from a wealthy family")

#
sjmisc::frq(elsoc_long_2016_2022.2$ahead_educ)
elsoc_long_2016_2022.2$ahead_educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ahead_educ, 
                      label = "Having a good level of education")

#
sjmisc::frq(elsoc_long_2016_2022.2$ahead_ambition)
elsoc_long_2016_2022.2$ahead_ambition <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ahead_ambition, 
                      label = "Having ambition")

#
sjmisc::frq(elsoc_long_2016_2022.2$ahead_hardwork)
elsoc_long_2016_2022.2$ahead_hardwork <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ahead_hardwork, 
                      label = "Hard work")

for (i in c("just_educ", "just_pension", "just_salud", "des_perc",
            "merit_effort","merit_talent",
            "ahead_family","ahead_educ","ahead_ambition", "ahead_hardwork")) {
elsoc_long_2016_2022.2[[i]] <-
sjlabelled::set_labels(x = elsoc_long_2016_2022.2[[i]],
                       labels = c("Strongly disagree","Disagree",
                                  "Neither disagree nor agree",
                                  "Agree", "Strongly agree"))
}

sjmisc::frq(elsoc_long_2016_2022.2$just_educ)
sjmisc::frq(elsoc_long_2016_2022.2$merit_effort)
sjmisc::frq(elsoc_long_2016_2022.2$merit_talent)
sjmisc::frq(elsoc_long_2016_2022.2$ahead_family)
sjmisc::frq(elsoc_long_2016_2022.2$ahead_educ)
sjmisc::frq(elsoc_long_2016_2022.2$ahead_ambition)
sjmisc::frq(elsoc_long_2016_2022.2$ahead_hardwork)
# Socioeconomic_________________________________________________________________

# Education_______________________________________
elsoc_long_2016_2022.2$educ <- 
  car::recode(elsoc_long_2016_2022.2$m01,
              "c(1,2,3)=1;c(4,5)=2;c(6,7)=3;c(8,9,10)=4; c(-888,-999)=NA")
elsoc_long_2016_2022.2$educ <-
  factor(elsoc_long_2016_2022.2$educ,
         labels = c("Primary","High school","Technical","Universitary"))

#reverse education, reference level is the highest level
#elsoc_long_2016_2022.2$educ <- forcats::fct_rev(elsoc_long_2016_2022.2$educ)

elsoc_long_2016_2022.2$educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$educ,
                      label = "Education")
sjmisc::frq(elsoc_long_2016_2022.2$educ)

#Recoding of education to years based on casen 2017.
elsoc_long_2016_2022.2$educyear<- as.numeric(
  car::recode(elsoc_long_2016_2022.2$m01, 
              "1=0;2=4.3;3=7.5;4=9.8;5=12.02;6=13.9;
               7=14.8;8=14.9;9=16.9;10=19.07;c(-888,-999)=NA", 
              as.numeric = T))

elsoc_long_2016_2022.2$educyear <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$educyear,
                      label = "Education in years")

class(elsoc_long_2016_2022.2$educyear)
sjmisc::frq(elsoc_long_2016_2022.2$educyear)

# Household income_________________________________________

#Impute midpoint of income ranges
elsoc_long_2016_2022.2$m30_rec <-
as.numeric(car::recode(elsoc_long_2016_2022.2$m30,
           "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=2700000;NA=NA;c(-888,-999)=NA"))

#Impute midpoint of income ranges (2021)
elsoc_long_2016_2022.2$m30b_rec <-
as.numeric(car::recode(elsoc_long_2016_2022.2$m30b,
           "1=125000;2=300000;3=400000;4=575000;5=70000;NA=NA;c(-888,-999)=NA"))

sjmisc::frq(elsoc_long_2016_2022.2$m30b_rec)

#Recode DK/DA of Income to NA
elsoc_long_2016_2022.2$m29_rec <-
  as.numeric(car::recode(elsoc_long_2016_2022.2$m29,"c(-888,-999)=NA"))

#replace NA of income with new imputed variable
elsoc_long_2016_2022.2$m29_imp <- 
  ifelse(test = !is.na(elsoc_long_2016_2022.2$m29_rec),
         yes =  elsoc_long_2016_2022.2$m29_rec,
         no =  elsoc_long_2016_2022.2$m30_rec)
summary(elsoc_long_2016_2022.2$m29_imp)

elsoc_long_2016_2022.2$m29_imp <- 
  ifelse(test = is.na(elsoc_long_2016_2022.2$m29_imp),
         yes =  elsoc_long_2016_2022.2$m30b_rec,
         no =  elsoc_long_2016_2022.2$m29_imp)
summary(elsoc_long_2016_2022.2$m29_imp)

# deflate at each year's prices
elsoc_long_2016_2022.2$deflactor <-
  with(elsoc_long_2016_2022.2, case_when(
    ola == 2016 ~ 113.88 / 123.82,
    ola == 2017 ~ 116.46 / 123.82,
    ola == 2018 ~ 119.45 / 123.82,
    ola == 2019 ~ 123.82 / 123.82
  ))

# N Household:
elsoc_long_2016_2022.2 <-
  elsoc_long_2016_2022.2 %>%
  mutate(n_hogar =
           dplyr::case_when(ola == 1 ~ nhogar1,
                            ola == 2 ~ m46_nhogar,
                            ola == 3 ~ m54,
                            ola == 4 ~ m54,
                            ola == 5 ~ m54))
sjmisc::frq(elsoc_long_2016_2022.2$n_hogar)

#Recode DK/DA to NA
elsoc_long_2016_2022.2$n_hogar_r<-
  car::recode(elsoc_long_2016_2022.2$n_hogar,"c(-888,-999)=NA")

# Per capita household income:
elsoc_long_2016_2022.2$ing_pc <- 
  (elsoc_long_2016_2022.2$m29_imp/elsoc_long_2016_2022.2$n_hogar_r)

elsoc_long_2016_2022.2$ing_pc <-
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ing_pc,
                      label = "Household income per capita")

sjmisc::descr(elsoc_long_2016_2022.2$ing_pc)

# Compute income quintiles
elsoc_long_2016_2022.2 <- elsoc_long_2016_2022.2 %>% 
  group_by(ola) %>% 
  mutate(quintil = ntile(-desc(ing_pc), 5)) %>% 
  ungroup()

elsoc_long_2016_2022.2$quintil <- 
  factor(elsoc_long_2016_2022.2$quintil,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

#reverse quintile, reference level is the highest quintile
#elsoc_long_2016_2022.2$quintil <- forcats::fct_rev(elsoc_long_2016_2022.2$quintil)

elsoc_long_2016_2022.2$quintil <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$quintil,
                      label = "Household income quintile per capita")

sjmisc::frq(elsoc_long_2016_2022.2$quintil)

#include new quintile category with missing cases
elsoc_long_2016_2022.2$quintil1<-
  car::recode(elsoc_long_2016_2022.2$quintil, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

#elsoc_long_2016_2022.2$quintil1 <- factor(elsoc_long_2016_2022.2$quintil1, c("Q1","Q2","Q3","Q4","Q5","QNA"))

elsoc_long_2016_2022.2$quintil1 <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$quintil1,
                      label = "Household income quintile per capita (NA)") 
sjmisc::frq(elsoc_long_2016_2022.2$quintil1)

#Age Ranges
elsoc_long_2016_2022.2$edad <- 
  factor(car::recode(elsoc_long_2016_2022.2$m0_edad, 
                     "18:29=1;30:49=2;50:64=3;65:150=4"),
         labels = c('18-29', '30-49', '50-64', '65 or more'))
elsoc_long_2016_2022.2$edad <-
  sjlabelled::set_label(elsoc_long_2016_2022.2$edad, 
                        label = c("Age groups")) 
# Other controls______________________________________________________________
# Subjective social status: individual ____________________________

elsoc_long_2016_2022.2$ess <- 
as.numeric(elsoc_long_2016_2022.2$d01_01)
sjmisc::frq(elsoc_long_2016_2022.2$ess)
#elsoc_long_2016_2022.2$ess <- sjmisc::rec(elsoc_long_2016_2022.2$ess,rec = "rev")
#sjmisc::frq(elsoc_long_2016_2022.2$ess)

elsoc_long_2016_2022.2$ess <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ess,
                      label = "Subjective Social Status: individual")

sjmisc::frq(elsoc_long_2016_2022.2$ess)
# sjPlot::plot_frq(elsoc_long_2016_2022.2$ess)

# Political position ____________________________
elsoc_long_2016_2022.2$pos_id <-
factor(
  car::recode(
    elsoc_long_2016_2022.2$c15,
    "c(11,12,-888,-999)='Does not identify';c(0,1,2,3,4)='Left';
     c(5)='Center';c(6,7,8,9,10)='Right'"
  ),
  levels = c('Left', 'Center', 'Right', 'Does not identify')
)

elsoc_long_2016_2022.2$pos_id <- factor(elsoc_long_2016_2022.2$pos_id,
                                      levels = levels(elsoc_long_2016_2022.2$pos_id))

elsoc_long_2016_2022.2$pos_id <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$pos_id, 
                      label = "Political identification")

sjmisc::frq(elsoc_long_2016_2022.2$pos_id)

# Select variables______________________________________________________________

df_study1 <- 
elsoc_long_2016_2022.2 %>% 
  filter(muestra == 1) %>% #keep original sample (vs refresh)
  select(idencuesta,ola,region,region_cod,comuna,comunacod=comuna_cod,
         muestra,
         sexo,
         # cuestion_mig,
         # starts_with("in_"),
         just_educ,
         just_pension,
         just_salud,
         des_perc,
         merit_effort,
         merit_talent,
         ahead_family,
         ahead_educ,
         ahead_ambition,
         ahead_hardwork,
         educ,educyear,
         # ing_pc,
         # quintil,
         quintil1,
         ess,
         # ess_inm,
         # nation_t1,
         # know_inm,
         # know_inm_bi,
         # frien_inm_bi_t4,
         # frien_inm_bi_t1,
         # know_inm_t1t4,
         # frien_inm,
         # frien_inm_bi,
         # frien_inm_t1t4,
         pos_id,
         edad)

# Reshape long to wide
df_study1_wide <- df_study1 %>% 
  tidyr::pivot_wider(id_cols = c("idencuesta","muestra"),
                     names_from = "ola",
                     # names_prefix = c("educ","educyear","ing_pc","quintil","quintil1"),
                     values_from = names(select(df_study1,comuna,comunacod,just_educ:edad,sexo))
                     )

# fix data to w01 values
df_study1_wide$edad_2 <-df_study1_wide$edad_1 #age
df_study1_wide$edad_3 <-df_study1_wide$edad_1
df_study1_wide$edad_4 <-df_study1_wide$edad_1
df_study1_wide$edad_5 <-df_study1_wide$edad_1
df_study1_wide$edad_6 <-df_study1_wide$edad_1

df_study1_wide$sexo_2 <-df_study1_wide$sexo_1 #sex
df_study1_wide$sexo_3 <-df_study1_wide$sexo_1
df_study1_wide$sexo_4 <-df_study1_wide$sexo_1
df_study1_wide$sexo_5 <-df_study1_wide$sexo_1
df_study1_wide$sexo_6 <-df_study1_wide$sexo_1

df_study1_wide$educ_2 <-df_study1_wide$educ_1 #education
df_study1_wide$educ_3 <-df_study1_wide$educ_1
df_study1_wide$educ_4 <-df_study1_wide$educ_1
df_study1_wide$educ_5 <-df_study1_wide$educ_1
df_study1_wide$educ_6 <-df_study1_wide$educ_1

df_study1_wide$educyear_2 <-df_study1_wide$educyear_1 #education years
df_study1_wide$educyear_3 <-df_study1_wide$educyear_1
df_study1_wide$educyear_4 <-df_study1_wide$educyear_1
df_study1_wide$educyear_5 <-df_study1_wide$educyear_1
df_study1_wide$educyear_6 <-df_study1_wide$educyear_1

df_study1_wide$quintil1_2 <-df_study1_wide$quintil1_1 #quintiles
df_study1_wide$quintil1_3 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_4 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_5 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_6 <-df_study1_wide$quintil1_1

df_study1_wide$ess_2 <-df_study1_wide$ess_1 # subjective status
df_study1_wide$ess_3 <-df_study1_wide$ess_1
df_study1_wide$ess_4 <-df_study1_wide$ess_1
df_study1_wide$ess_5 <-df_study1_wide$ess_1
df_study1_wide$ess_6 <-df_study1_wide$ess_1

df_study1_wide$pos_id_2 <-df_study1_wide$pos_id_1 # political position
df_study1_wide$pos_id_3 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_4 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_5 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_6 <-df_study1_wide$pos_id_1

df_study1_wide$comuna_2 <-df_study1_wide$comuna_1 #comuna
df_study1_wide$comuna_3 <-df_study1_wide$comuna_1
df_study1_wide$comuna_4 <-df_study1_wide$comuna_1
df_study1_wide$comuna_5 <-df_study1_wide$comuna_1
df_study1_wide$comuna_6 <-df_study1_wide$comuna_1

df_study1_wide$comunacod_2 <-df_study1_wide$comunacod_1 #comuna
df_study1_wide$comunacod_3 <-df_study1_wide$comunacod_1
df_study1_wide$comunacod_4 <-df_study1_wide$comunacod_1
df_study1_wide$comunacod_5 <-df_study1_wide$comunacod_1
df_study1_wide$comunacod_6 <-df_study1_wide$comunacod_1

dim(df_study1_wide)

sjPlot::view_df(df_study1_wide,
                show.frq = T,show.values = T,show.na = T,show.prc = T, show.type = T)

# reshape from long to wide
pacman::p_load(datasets,data.table)
df_study1_long <- data.table::melt.data.table(data.table::setDT(df_study1_wide),
              id.vars = c("idencuesta","muestra"),
              variable.name = c("ola"),
              measure = patterns("^comuna_","^comunacod_",
                                 "^just_educ_", "^just_pension_", "^just_salud_", "^des_perc_",
                                 "^merit_effort_","^merit_talent_",
                                 "^ahead_family_", "^ahead_educ_", "^ahead_ambition_", "^ahead_hardwork_",
                                 "^educ_","^educyear_","^quintil1_","^ess_",
                                 "^pos_id_",
                                 "^edad_","^sexo_"
                                 ),
              value.name = c("comuna","comunacod",
                             "just_educ", "just_pension","just_salud", "des_perc",
                             "merit_effort","merit_talent",
                             "ahead_family","ahead_educ","ahead_ambition","ahead_hardwork",
                             "educ","educyear","quintil1","ess",
                             "pos_id",
                             "edad","sexo"),
              na.rm = F,value.factor = TRUE
              )

names(df_study1_long) #check names of long dataset
dim(df_study1_long) #check dimensions of the dataframe
# Original dataset with 6 waves
df_study2_long <- df_study1_long

# filter the dataset for the waves 1 to 4 and 6
df_study1_long <-
df_study1_long %>% 
  filter(ola %in% c(1,2,3,4,6)) %>% 
  mutate(ola=factor(ola,levels = 1:6,labels = 1:6))
dim(df_study1_long) #check, now is OK

df_study1_long <- 
set_label(x = df_study1_long,
          label = get_label(select(df_study1,names(df_study1_long))))

sjPlot::view_df(df_study1_long,
                show.frq = T,show.values = T,show.na = T,show.prc = T,
                show.type = T)
#______________________________________________________________________________
# obtain the idencuesta for wave 6
ids <- 
  df_study1 %>% 
  select(idencuesta,ola) %>% 
  filter(ola==6) %>% 
  sjmisc::frq(idencuesta,show.na = F) %>% as.data.frame()

# filter data by the idencuesta of t5
df_study1_long_t6 <- 
  df_study1_long %>%
  filter(idencuesta %in% ids$val)

names(df_study1_long_t6)
dim(df_study1_long_t6)
sjmisc::frq(df_study1_long_t6$ola)

# SAVE DATA____________________________________________________________________
save(df_study1_long,file = here::here("input/data-proc/df_study1_long.RData"))
save(df_study1_long_t6,file = here::here("input/data-proc/df_study1_long_t6.RData"))
save(df_study2_long,file = here::here("input/data-proc/df_study2_long.RData"))
# save codebook
sjPlot::view_df(df_study1_long,
        show.frq = T,
        show.prc = T,
        show.na = T,
        file = here::here("output/df_study1_long.html"))

sjPlot::view_df(df_study1_long_t6,
        show.frq = T,
        show.prc = T,
        show.na = T,
        file = here::here("output/df_study1_long_t6.html"))
```

## Contextual data

Notas de Trabajo de campo:

* CENSO 2017, día 19 de abril de 2017. (pegar a 2016)
* CASEN 2017, 2 Noviembre 2017 – 4 Febrero 2018. (pegar a 2017)
* CASEN 2020, 31 de octubre 2020 – 04 de febrero 2021. (pegar a 2020)

* Faltarían datos 2019 para algunas comunas

![](https://radiografia-cambio-social-2016-2021.netlify.app/reporte-elsoc_files/figure-html/amen1-wave-1.png)

```{r}
remove(list = ls()) # clean workspace
# .rs.restartR() #restart R session
if (!require("pacman")) install.packages("pacman") # install pacman
# load libraries
pacman::p_load(dplyr,readxl,sjmisc, sjlabelled, questionr, car,here)
load(here::here("input/data-original/cit17.Rdata"))
load(here::here("input/data-original/cit18.Rdata"))
load(here::here("input/data-original/cit19.Rdata"))
load(here::here("input/data-original/casen2017.Rdata"))
load(here::here("input/data-original/casen2022.Rdata"))
load(here::here("input/data-original/censo17.Rdata"))
load(here::here("input/data-original/elsoc_long.Rdata"))

df_ine <- 
readxl::read_excel("input/data-original/data_comunas_ine_dem.xlsx")
df_ine$comuna_lc <- tolower(df_ine$comuna)

df_pro_ine <- 
readxl::read_excel("input/data-original/estimaciones-y-proyecciones-2002-2035-comunas.xlsx")
names(df_pro_ine)
df_proye <- 
df_pro_ine %>% 
  select(comuna_cod=Comuna,comuna_nom=`Nombre Comuna`,
         "sexo"=Sexo_1_Hombre_2_Mujer,
         "edad"=Edad,
         pob_17=`Poblacion 2017`,
         pob_18=`Poblacion 2018`,
         pob_19=`Poblacion 2019`,
         pob_20=`Poblacion 2020`)
names(df_proye)

# calculate population proyections 
df_proye_comunas <- 
df_proye %>% 
  group_by(comuna_cod, comuna_nom) %>% 
  summarise(pob_17=sum(pob_17),
            pob_18=sum(pob_18),
            pob_19=sum(pob_19),
            pob_20=sum(pob_20))

df_proye_comunas <-
df_proye_comunas %>% 
  mutate(pob_total_17=sum(df_proye_comunas$pob_17),
         pob_total_18=sum(df_proye_comunas$pob_18),
         pob_total_19=sum(df_proye_comunas$pob_19),
         pob_total_20=sum(df_proye_comunas$pob_20))
names(df_proye_comunas)
# ELSOC long data ____________________________________________________________
df_elsoc <- 
elsoc_long_2016_2022.2 %>% 
  select(comuna,comuna_cod) %>%
  mutate(comuna_lc=tolower(comuna)) %>% 
  group_by(comuna,comuna_lc,comuna_cod) %>% 
  summarise(n_elsoc=n())

df_a <- full_join(df_elsoc,df_ine, by = "comuna_lc", suffix = c("_elsoc","_ine"))
summary(df_a)

# ELSOC CIT - territorial data _________________________________________________
df_17 <- 
cit17 %>% 
  select(idencuesta,escolaridad)

df_18 <- 
cit18 %>% 
  select(idencuesta,escolaridad)

df_19 <- 
cit19 %>% 
  select(idencuesta,escolaridad)
df_19[df_19==-995] <- NA

# create data set
df_cit_a <- 
elsoc_long_2016_2022.2 %>% 
  select(comuna,comuna_cod,idencuesta) %>% 
  left_join(x = .,y = df_19, by ="idencuesta") %>% 
  select(-idencuesta)
#check for duplicates
df_cit_b <- 
  df_cit_a %>% distinct() %>% arrange(comuna_cod, escolaridad)

#generate final dataset
df_cit <- 
  df_cit_b %>% 
  group_by(comuna, comuna_cod) %>% 
  summarise(esc_cit17=mean(escolaridad,na.rm = T))

sjmisc::frq(df_cit$esc_cit17)

# Select and organize variables________________________________________________
df_study1_comunas <- 
df_cit %>%
  mutate(comuna=sjlabelled::as_character(comuna_cod)) %>% 
  dplyr::select(comuna,
                comunacod=comuna_cod,
                esc_cit17
                )
# names(df_study1_comunas)

# Label data____________________________________________________________________
sjlabelled::set_label(df_study1_comunas) <- 
c("Comuna",
  "Codigo Comuna",
  "Escolarity average - CIT 2017 - Municipality (using census tract)"
)

# save data____________________________________________________________________
save(df_study1_comunas,file = here::here("input/data-proc/study_1_comunas.RData"))

# save codebook
sjPlot::view_df(df_study1_comunas,
        show.frq = T, 
        show.prc = T, 
        show.na = T,
        file = here::here("output/df_study1_comunas.html"))
```

# Merge individual & contextual

```{r}
remove(list = ls()) # clean workspace
# .rs.restartR() #restart R session
if (!require("pacman")) install.packages("pacman") # install pacman
# load libraries
pacman::p_load(dplyr,readxl,sjmisc, sjlabelled, questionr, car,here)

load(file = here::here("input/data-proc/df_study1_long.RData"))
load(file = here::here("input/data-proc/df_study2_long.RData"))
load(file = here::here("input/data-proc/df_study1_long_t5.RData"))
load(file = here::here("input/data-proc/study_1_comunas.RData"))
load(file = here::here("input/data-proc/df_gini_comunas_wide.RData"))
load(file = here::here("input/data-proc/df_gini_comunas_long.RData"))
df_gini_comunas_wide <-  df_gini_comunas_wide %>% dplyr::rename(comunacod=COMUNA_COD) 
df_gini_comunas_wide$comunacod <- as.numeric(df_gini_comunas_wide$comunacod)

df_study1_long_comunas <- 
 dplyr::left_join(x = df_study1_long,
                  y = df_study1_comunas, 
                  by = "comunacod") %>% 
 dplyr::select(idencuesta:edad,sexo,
               esc_cit17) %>% 
 dplyr::rename(comuna=comuna.x)

df_study1_long_comunas <- 
 dplyr::left_join(x = df_study1_long_comunas,
                  y = select(df_gini_comunas_wide, comunacod,starts_with("gini")), 
                  by = c("comunacod"))

df_study2_long_comunas <- 
 dplyr::left_join(x = df_study2_long,
                  y = df_study1_comunas, 
                  by = "comunacod") %>% 
 dplyr::select(idencuesta:edad,sexo,
               esc_cit17) %>% 
 dplyr::rename(comuna=comuna.x)

df_study2_long_comunas <- 
 dplyr::left_join(x = df_study2_long_comunas,
                  y = select(df_gini_comunas_wide, comunacod,starts_with("gini")), 
                  by = c("comunacod"))

df_study_long_t5_comunas <- 
 dplyr::left_join(x = df_study1_long_t5,
                  y = df_study1_comunas, 
                  by = "comunacod") %>% 
 dplyr::select(idencuesta:edad,sexo,
               esc_cit17) %>% 
 dplyr::rename(comuna=comuna.x)

# save data____________________________________________________________________
save(df_study1_long_comunas,
     file = here::here("input/data-proc/df_study1_long_comunas.RData"))
save(df_study2_long_comunas,
     file = here::here("input/data-proc/df_study2_long_comunas.RData"))
save(df_study_long_t5_comunas,
     file = here::here("input/data-proc/df_study_t5_ind_comunas.RData"))

# save codebook
sjPlot::view_df(df_study1_long_comunas,
        show.frq = T, 
        show.prc = T, 
        show.na = T,
        file = here::here("output/df_study1_long_comunas.html"))

sjPlot::view_df(df_study_long_t5_comunas,
        show.frq = T, 
        show.prc = T, 
        show.na = T,
        file = here::here("output/df_study_long_t5_comunas.html"))
```