### 2024-06-25 presentations/ISPP-santiago2024/ispp2024.qmd
up ispp presentation

---
pagetitle: "Changes in the justification of educational inequalities: The role of perceptions of inequality and meritocracy during the COVID pandemic"
author: Juan Carlos Castillo, Julio Iturra & Kevin Carrasco
bibliography: ../../input/bib/merit-edjust.bib 
csl: ../../input/bib/apa7-singlespace.csl

knitr:
  opts_chunk: 
    collapse: true
    comment: "#>" 
    R.options:
      knitr.graphics.auto_pdf: true

format: 
  revealjs:
    logo: images/edumer.png
    slide-number: true
    theme: [pp.scss]
    #theme: [pp.scss]
    title-slide-attributes:
      visibility: false
#      data-background-image: images/portada.jpg
#      data-background-size: cover
    auto-play-media: true

editor: visual
---

```{r data, results=FALSE, warning=FALSE, echo=FALSE}
if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               summarytools,
               ggplot2,
               sjPlot,
               tidyverse,
               ggalluvial,
               survey,
               shadowtext,
               srvyr
               )
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))

```

::: {.columns .v-center-container}
::: {.column width="30%"}
![](qr.png){width="80%" fig-align="right"}
<br>
![](images/logocoes.png){width="90%" fig-align="right"}
<br>
![](images/edumer.png){width="80%" fig-align="left"}
:::

::: {.column width="70%"}

## **Changes in the justification of educational inequalities:**

### The role of perceptions of inequality and meritocracy during the COVID pandemic

----------------------------------------------------------------

Juan Carlos Castillo, Julio Iturra & **Kevin Carrasco* **

#### Fondecyt N° 1210847: Meritocracia en la escuela
#### Center for social conflict and cohesion studies

**ISPP Annual meeting** Chile, July 2024
:::
:::

# Introduction {data-background-color="black"}

## Introduction

::: {.incremental}

::: {.columns .v-center-container}
::: {.column width="50%" .incremental}

### Justification of inequality

- Perceived economic inequality and justification of inequality [@wegenerIllusionDistributiveJustice1987; @jassoJusticeEarningsNew1978; @janmaatSubjectiveInequalityReview2013]

- Cultural factors and policy feedback effects [@kelley_legitimate_2021; @mijs_paradox_2021; @trump_income_2018]

- Chile, marketization & neoliberal reforms [@bellei_estudio_2013; @corvalanMercadoEscolarLibertad2016]

:::

::: {.column width="50%" .incremental}
### Meritocracy

- Effort and talent [@young_rise_1962]

- Normative and perceptual elements [@duru-bellat_who_2012; @castillo_meritocracia_2019]

:::
:::

::: {.r-stack style="font-size: 130%;"}
- Covid pandemic [@breznau_welfare_2021; @hankivsky_introduction_2022]
:::

:::

## This study

<br>

![](general-scheme.png){width="100%" fig-align="center"}

# Methods {data-background-color="black"}

## Data

::: {.incremental}

- Longitudinal social study of Chile (ELSOC-COES) 2016-2022

- 2021 (telephone version) the dependent variable was not asked

- N total=8042; 1689 individuals for 5 waves

:::

## Variables

```{r bivariate-plots, echo=FALSE}
df_study1 <- df_study_long_t6_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
merit_plot <- df_study1 %>% 
  select(merit_effort, merit_talent, des_perc, just_educ) %>% 
  sjPlot::plot_stackfrq(weight.by = df_study1$ponderador_long_total,
                        expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      vjust=rep(c("bottom", "top", "bottom", "top", "bottom"),4),
                      legend.labels = c("Strongly disagree", "Disagree", "Neither agree
nor disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Merit effort", "Merit talent", "Inequality 
perception", "Justification 
of 
educational inequality")
                      ) +
  theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1)) +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2022 (n = 8042)")
merit_plot

```

## Variables {.smaller}

```{r error=FALSE}
df_study1 <- df_study_long_t6_comunas %>% dplyr::select(ola,
                                               just_educ, des_perc, merit_talent, merit_effort,
                                               educ, quintil1:sexo)

df_study1$ola <- factor(df_study1$ola,
         labels = c("2016","2017","2018","2019", "2022"))
df_study1$ola <- 
sjlabelled::set_label(x = df_study1$ola,
                      label = "Wave")

df_study1$sexo <- factor(df_study1$sexo,
         labels = c("Male", "Female"))
df_study1$sexo <- 
sjlabelled::set_label(x = df_study1$sexo,
                      label = "Gender")

for (i in c("just_educ", "des_perc",
            "merit_effort","merit_talent")) {
  df_study1[[i]] <- factor(df_study1[[i]],
                            levels = 1:5,
                            labels = c("Strongly disagree","Disagree",
                                       "Neither disagree nor agree",
                                       "Agree", "Strongly agree"))
}

df_study1$just_educ <- 
sjlabelled::set_label(x = df_study1$just_educ, 
                      label = "Justification of educational inequality")

#
df_study1$des_perc <- 
sjlabelled::set_label(x = df_study1$des_perc, 
                      label = "Income differences are too large")

#
df_study1$merit_effort <- 
sjlabelled::set_label(x = df_study1$merit_effort, 
                      label = "People are rewarded for their efforts")

#
df_study1$merit_talent <- 
sjlabelled::set_label(x = df_study1$merit_talent, 
                      label = "People are rewarded for their intelligence and skills")

lab_study1 <- sjlabelled::get_label(df_study1) 
df_study1 <- df_study1 %>% na.omit() %>% sjlabelled::set_label(lab_study1)
```

::: {.columns .v-center-container}

::: {.column width="50%"}

```{r error=FALSE}
df<- dfSummary(x = select(df_study1, ola, quintil1, sexo),
               plain.ascii = FALSE,
               style = "grid",
               tmp.img.dir = "/tmp",
               graph.magnif = TRUE,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = F, # n valido
               col.widths = c(10,10,10)
               )
df$Variable <- NULL # delete variable column

print(df, method='render',
      footnote = NA, 
      headings = FALSE, 
      style = list(
        table = "font-size: 8px;",
        th = "font-size: 8px;",
        td = "font-size: 8px;"
      ))
```

:::

::: {.column width="50%"}
```{r error=FALSE}
df<- dfSummary(x = select(df_study1, edad, pos_id),
               plain.ascii = FALSE,
               style = "grid",
               tmp.img.dir = "/tmp",
               graph.magnif = TRUE,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = F, # n valido
               col.widths = c(10,10,10)
               )
df$Variable <- NULL # delete variable column

print(df, method='render',
      footnote = NA, 
      headings = FALSE, 
      style = list(
        table = "font-size: 8px;",
        th = "font-size: 8px;",
        td = "font-size: 8px;"
      ))
```
:::

:::

# Results {data-background-color="black"}

##

### Changes in the justification of educational inequality (2016-2022)

```{r}
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t6_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
df_study1 <- df_study1[complete.cases(df_study1$segmento), ]

df_study1$just_educ <- factor(df_study1$just_educ,
         labels = c("Strongly 
disagree", "Disagree", "Neither agree
nor disagree", "Agree", "Strongly 
agree"))

elsoc_diseno <- svydesign(ids = ~segmento, #muestreo por conglomerado a nivel de manzanas (segmento)
                          strata = ~estrato, #muestreo estratificado a nivel ciudad (estato)
                          weights = ~ponderador_long_total, #ponderador longitudinal
                          nest = TRUE,
                          data = df_study1)

#Paso 1
datos.justeduc <- data.frame((svytable(~just_educ + ola + idencuesta, elsoc_diseno, round = F))) %>% dplyr::filter(Freq>0)  %>% group_by(ola) %>% mutate(porcentaje=Freq/sum(Freq)) %>% na.omit()

#Paso 2
etiquetas.justeduc <- data.frame((svytable(~just_educ + ola, elsoc_diseno, round = F))) %>% group_by(ola) %>% mutate(porcentaje=Freq/sum(Freq)) %>% na.omit() %>% 
  mutate(idencuesta = 1)
datos.justeduc$just_educ <- factor(datos.justeduc$just_educ, levels = rev(levels(datos.justeduc$just_educ)))
etiquetas.justeduc$just_educ <- factor(etiquetas.justeduc$just_educ, levels = rev(levels(etiquetas.justeduc$just_educ)))
```

```{r alluvial}
colors<- c("#f1eef6ff","#bdc9e1ff","#b3b3b3ff","#74a9cfff","#0570b0ff")
alluvial_justeduc <- ggplot(datos.justeduc, aes(x = ola, fill = just_educ, stratum = just_educ,
                             alluvium = idencuesta, y = porcentaje))+
    ggalluvial::geom_flow(alpha = .66) + 
    ggalluvial::geom_stratum(linetype = 0) +
    scale_y_continuous(labels = scales::percent) + 
    ylab(label = NULL) +
    xlab(label = NULL) + 
    theme(#legend.position = 'top',
          legend.title = element_blank(),
          plot.caption = element_text(hjust = 1)) +
    scale_fill_manual(values = colors) +
    geom_shadowtext(data = etiquetas.justeduc, 
              aes(label = ifelse(porcentaje > 0 , scales::percent(porcentaje, accuracy = .1),"")),
              position = position_stack(vjust = .5),
              show.legend = FALSE,
              size = 4,
              color = rep('white'),
              bg.colour='grey30')+
  scale_x_discrete(labels = c(2016, 2017, 2018, 2019, 2022)) +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2022 (n = 8042)")

alluvial_justeduc
```

##

### Changes in the main independent variables over years (2016-2022)

```{r}
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t6_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
df_study1 <- df_study1[complete.cases(df_study1$segmento), ]

elsoc_diseno <- as_survey_design(df_study1,
                                 ids = segmento, #muestreo por conglomerado a nivel de manzanas (segmento)
                          strata = estrato, #muestreo estratificado a nivel ciudad (estato)
                          weights = ponderador_long_total, #ponderador longitudinal
                          nest = TRUE
                          )

just_years_mean = elsoc_diseno %>% #filter(!is.na(ola)) %>% 
  group_by(ola) %>%
  summarise(education = survey_mean(just_educ, vartype = "ci", na.rm = T)) %>% 
  ungroup()

just_years <- just_years_mean %>% 
  pivot_longer(cols = c("education"), 
               names_to = "variable",
               values_to = "value") %>% 
  cbind(type=c("Dependent")) %>% 
  select(ola, variable, value, type, value_upp=education_upp, value_low=education_low)

ineq_years = elsoc_diseno %>% #filter(!is.na(ola)) %>% 
  group_by(ola) %>%
  summarise(inequality = survey_mean(des_perc, vartype = "ci", na.rm = T),
            ) %>% 
  ungroup()

ineq_years <- ineq_years %>% 
  pivot_longer(cols = c(
                        "inequality"), 
               names_to = "variable",
               values_to = "value") %>% 
  cbind(type=c("Independent")) %>% 
  select(ola, variable, value, type, value_upp=inequality_upp, value_low=inequality_low)

talent_years = elsoc_diseno %>% #filter(!is.na(ola)) %>% 
  group_by(ola) %>%
  summarise(
            talent = survey_mean(merit_talent, vartype = "ci", na.rm = T)) %>% 
  ungroup()

talent_years <- talent_years %>% 
  pivot_longer(cols = c(
                        "talent"), 
               names_to = "variable",
               values_to = "value") %>% 
  cbind(type=c("Independent")) %>% 
  select(ola, variable, value, type, value_upp=talent_upp, value_low=talent_low)

effort_years = elsoc_diseno %>% #filter(!is.na(ola)) %>% 
  group_by(ola) %>%
  summarise(
            effort = survey_mean(merit_effort, vartype = "ci", na.rm = T)) %>% 
  ungroup()

effort_years <- effort_years %>% 
  pivot_longer(cols = c(
                        "effort"), 
               names_to = "variable",
               values_to = "value") %>% 
  cbind(type=c("Independent")) %>% 
  select(ola, variable, value, type, value_upp=effort_upp, value_low=effort_low)

vars_years <- rbind(just_years, ineq_years, talent_years, effort_years)

categorias_x <- c(2016, 2017, 2018, 2019, 2022)
vars_years$variable <- factor(vars_years$variable, levels = c("education", "inequality", "talent", "effort"), labels=  c("Justification
educational
inequality
(dependent)", "Inequality
perception", "Merit: Talent", "Merit: Effort"),
ordered = TRUE)

categorias_z <- c("Justification
educational
inequality
(dependent)", "Inequality
perception
(independent)", "Merit: Talent
(independent)", "Merit: Effort
(independent)")

## Crear el gráfico principal
years_plot <- ggplot(vars_years, aes(ola, value, color = variable, shape = variable, linetype = variable)) +
  theme_bw(base_size = 14) +
  geom_point(size = 4) +
  geom_line(aes(group = variable)) +
  geom_errorbar(aes(ymin = value_low, ymax = value_upp), width = 0.2) +
  ylab("Mean") +
  xlab("Wave") +
  ylim(1, 5) +
  scale_x_discrete(labels = categorias_x) +
  scale_color_manual(labels = categorias_z, name = "Variable", values =  c("#bdc9e1ff","#0570b0ff","#74a9cfff","#b3b3b3ff")) +
  scale_shape_manual(labels = categorias_z, name = "Variable", values = c("square", "circle", "triangle", "diamond")) +
  scale_linetype_manual(labels = categorias_z, name = "Variable", 
                        values = c("solid", "dashed", "dotted", "dotdash")) +  # Definir diferentes tipos de línea
   theme(legend.position = "none") +
  coord_cartesian(xlim = c(1, 6)) +  # Ampliar el eje x
  scale_x_discrete(labels = c(2016, 2017, 2018, 2019, 2022, ""))  # Personalizar etiquetas del eje x

# Obtener los datos para etiquetar las líneas
label_data <- vars_years %>%
  group_by(variable) %>%
  filter(row_number() == n()) %>%
  ungroup() %>%
  mutate(label_x = 5.5, label_y = value)

# Añadir etiquetas personalizadas
years_plot <- years_plot +
  geom_point(data = label_data, aes(label_x - 0.05, label_y, color = variable, shape = variable), size = 4) +  # Capa de puntos para los símbolos de forma
  geom_text(data = label_data, aes(label_x, label_y, label = variable), color = "black", hjust = 0, nudge_x = 0.05, nudge_y = 0, size = 4, lineheight = 0.8) +  # Ajustar el interlineado del texto de las etiquetas
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2022 (n = 8042)")
years_plot
```

##

### Multivariate analysis

```{r distribution-educ}
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               summarytools,
               lme4,
               texreg
               )
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))

# generate analytical sample
df_study1 <- 
  df_study_long_t6_comunas %>%
  select(idencuesta,ola,comuna, ponderador_long_total,just_educ, des_perc, merit_effort, merit_talent, education=educ,quintil1:sexo) %>% 
  na.omit() %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2,
         sexo=as.factor(sexo)
         )

h6 <- "ola+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent+merit_effort"
educ6 <- lmer(formula(paste0("just_educ~",h6,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)

plot_model(educ6, 
            rm.terms = c("educationHigh school", "educationTechnical", "educationUniversitary", "ess", "quintil1Q5"), #quitar terminos
            title = "", #quitar titulo
            show.values = TRUE, #mostrar valor de efectos
            dot.size = 1.5, #tamaño circulos
            line.size = 1, #tamaño CI
            value.size = 3.5, #tamaño valor efectoss
            spacing = 1, #espacio entre efectos
            vline.color = "red", # linea roja en punto neutro (1)
            axis.labels = rev(c("Year 2017", "Year 2018", "Year 2019", "Year 2022", 
                                "Quintile 2", "Quintile 3", "Quintile 4", "Quintile 5",
                                "Age 30-49", "Age 50-64", "Age 65 or more",
                                "Female", "Center", "Right", "Does not identify",
                                "Inequality perception", "Merit: Talent", "Merit: Effort")), # con rev porque automatico los tira en otro orden
            group.terms = c(1,1,1,2,3,3,3,3,4,4,4,1,5,5,5,1,2,2),
            value.offset = 0.4  # Ajusta este valor para mover el texto de los coeficientes hacia arriba
) + # variables dependientes
  theme_bw(base_size = 16)
```

# Conclusions {data-background-color="black"}

## Conclusions

::: {.incremental}

- Decreasing pattern of justification of educational inequality between 2017-2019, but a increase in the last measurement point (2022).

- Mixed evidence about the association between the perception of meritocracy and justification of inequality across time.

- Throughout the years, a larger perceived economic inequality motivates a lower justification of inequality in education

:::

## Future research

::: {.incremental}

- Access vs outcomes in education

- 7 wave of ELSOC (2024)

- Justification of inequality in pensions and health (market justice preferences)
:::

# Thank you!

- Kevin Carrasco ([kevin.carrasco@ug.uchile.cl](kevin.carrasco@ug.uchile.cl))

- Github Repo: [https://github.com/justicia-distributiva/merit-edjust](https://github.com/justicia-distributiva/merit-edjust)

# Anexos {data-background-color="black"}

## Interaction effects

```{r}
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               summarytools,
               lme4,
               texreg,
               ggplot2
               )
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))

# generate analytical sample
df_study1 <- 
  df_study_long_t6_comunas %>%
  select(idencuesta,ola,comuna, ponderador_long_total,just_educ, des_perc, merit_effort, merit_talent,education=educ,quintil1:sexo) %>% 
  na.omit() %>% 
  mutate(ola = as.numeric(ola),
         ola_2=as.numeric(ola)^2,
         sexo=as.factor(sexo)
 #        edad=as.numeric(edad)
         #pos_id=as.numeric(pos_id)
         )
h1b <- "ola+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent+merit_effort+merit_effort*ola"
educ1 <- lmer(formula(paste0("just_educ~",h1b,"+(ola+merit_effort|idencuesta)")),data = df_study1, weights = ponderador_long_total)

plot_model(educ1, 
            type="int"
) + # variables dependientes
  theme_bw(base_size = 16)+
  labs(title="",
       x="Wave",
       y="Justification of educational inequality",
       color="Merit: Effort")
```

## References