### 2024-12-30 02-antecedentes.qmd

# The justification of educational inequalities

The justification of economic inequalities is a controversial field of study. As equality is usually considered a synonym of justice in common sense, the idea that someone could be willing to justify inequalities seems (in principle) far from what is considerably reasonable. Nevertheless, and as pointed out by Barry, the whole normative discussion in social justice can be conceived as “the defensibility of unequal relations between people” [-@barry_theories_1989, p. 3], whereby empirical research has consistently shown strong variations between individuals and societies regarding inequality justification [@trump_income_2018;@kelley_legitimate_2021;@mijs_paradox_2021]. Both theoretical and empirical approaches have been able to establish a clear distinction between inequality and justice, opening the question of to what extent inequality can be tolerated or even justified.

Studies addressing inequality justification have assessed the support for market ideologies, salary gaps, and distributive preferences. A strand of this research has focused on specific policy areas such as health, pensions, and education,  attending to the degree of justification of market-based access to social services. For instance, @lindhPublicOpinionMarkets2015 found that public support for the market distribution of social services is generally low in most countries, implying that most people believe it is unfair for market forces to determine access to basic social services. Similarly, @busemeyer_welfare_2020 shows that high-income citizens are less supportive of expanding public spending on universal benefit schemes. Instead, they tend to prefer a public basic insurance scheme, even if it provides relatively more benefits to low-income individuals.

In the area of empirical studies, the operationalization of educational inequality justification in terms of questionnaire items has been diverse, and there is not a single instrument or scale dedicated specifically to this concept. Rather, some attitudinal surveys usually include education-related items that are posteriorly used as dependent or independent variables. Some examples are the International Social Survey Programme (ISSP) (_Is it just or unjust – right or wrong – that people with higher incomes can buy better education for their children than people with lower incomes?_), and the European Social Survey (ESS) 2018 (_Compared to other people in [country], I have had a fair chance of achieving the level of education I was seeking_). Although both questions have been used to measure the justification of educational inequality, they refer to different aspects: the ESS can be classified as reflexive, as the respondent is involved as a referent for the question, whereas the ISSP question is non-reflexive. Furthermore, ISSP refers to educational access, while the ESS refers to educational outputs.

_Meritocracy and educational inequalities_

The relationship between performance and outputs brings about to the concepts of merit and meritocracy, key in educational justice research. The concept of meritocracy, coined by the sociologist Michael Young in his satirical work "The Rise of the Meritocracy" [-@young_rise_1962], has a rich history and multifaceted meaning. Young's term was initially used to critique a hypothetical future society where social and economic positions were entirely determined by individuals' merits, talents, and efforts, ostensibly eliminating the influence of inherited privilege. However, over time, the term has evolved to describe a social system where advancement and success are perceived as directly linked to one's abilities and achievements, sidelining inequalities of opportunities [@bell_politics_2020]. This would contribute to the justification of inequality in the educational domain [@sandel_tyranny_2020], as individual ability is often inherited, and educational attainment is influenced by family background [@goldin_meritocracy_2000].

Although there are few studies relating meritocratic beliefs and inequality justification, from a broader perspective it is possible to find related research which incorporates the support of different distributive principles with regard to education. For instance, in an experiment by @igliozziFairShareEffects2024, they establish how the support for different justice principles (equity, equality, and need) is related to support for redistributive policies in domains such as education. They find that a larger system justification  [@jost_psychology_2003] is related to higher support for the principles of equity and equality (over need) in distributing educational outcomes, where equity was measured in terms of access to education related to performance ('Students who place within the highest 50% of testing scores from the previous year can attend the magnet school'). In the same line, @leeFairnessPerceptionsEducational2023 show that performance assessments (in what they call “neoliberal orientations”) predicted people's fair perceptions of educational inequality.

Empirical cross-sectional evidence from Chile reveals two significant trends: low public support for the perception of meritocracy and limited justification for unequal access to social services such as education and health. However, the relationship between meritocracy and the justification of inequality within the educational domain remains underexplored. For instance, @castillo_clivajes_2013 analyzed data from the International Social Survey Programme (ISSP) in 1999 and 2009, highlighting consistently low support for inequality based on resource allocation or individual merit. Similarly, @Castilloetal2019 demonstrated widespread skepticism among Chileans regarding the meritocratic system, with many disagreeing with the notion that rewards are fairly distributed based on talent and effort.

More specific empirical research examining the relationship between meritocratic perceptions and the justification of educational inequality provides valuable insights into the dynamics shaping individuals' attitudes and beliefs. Some studies have explored the extent to which people believe that educational systems are meritocratic and how these perceptions influence their acceptance or rejection of educational disparities, as well as status differences [@McCoy2007]. This belief is particularly influential in educational institutions, where it not only encourages academic success but also legitimizes social and income inequality [@batruch_belief_2023]. Most of the evidence so far suggests that those who believe in a direct link between effort, talent, and educational success are more inclined to rationalize existing disparities in educational outcomes. However, empirical findings also reveal nuances in this relationship. For instance, studies show that when individuals perceive structural barriers, such as economic disparities or discrimination, as hindering the meritocratic ideal, they may be less likely to justify educational inequalities [@dayPerceivedIdealInequality2023]. This suggests that while meritocratic beliefs play a significant role, individuals' awareness of broader societal issues can mediate their willingness to rationalize educational disparities.

_This study_

The present research aims at identifying changes in the justification of educational inequalities in Chile, with a particular focus in its relationship with meritocratic perceptions. The time frame for this analysis covers between 2016 and 2023, by using longitudinal panel survey data provided by the ELSOC database (Estudio Longitudinal Social de Chile). Although this period comprises the COVID pandemic, the changes in the variables analyzed are not possible to be attributed only to this event in terms of a health crisis, as several economic, social and cultural processes overlapped worldwide. Besides, in the specific case of Chile there were two constitutional processes happening in this very same period after a massive social upheaval.

As there are measures before and after the pandemic period, firstly we explore to what extent it is possible to identify variations in inequality justification within this time frame. Although we are aware that it is not possible to make a causal claim given that this period is characterized by many other circumstances besides COVID, we argue that the increase of risk perceptions in critical situations (such as the pandemic) impacts public concerns about equality and redistribution [@breznau_welfare_2021].

Within this framework, we point out that periods of rising socioeconomic and health vulnerabilities could weaken the justification of social inequalities. This change would be at least in part linked to a challenge of the meritocratic ideals. Chile provides a significant case for studying inequality, as it exemplifies a prosperous yet highly unequal society within Latin America. Under such conditions, a shock like the COVID-19 pandemic posed a considerable threat to economically vulnerable groups and those in precarious labor market positions. Chile implemented sanitary controls early in the pandemic, initiating lockdowns on March 18, 2020, and lifting all restrictions by March 5, 2023. In this context, Chile stands out as one of the strictest countries globally and within South America in terms of pandemic-related policies. According to the Stringency Index [@hale_global_2021]—ranging from 0 to 100 (100 being the strictest) and encompassing measures like school closures, workplace shutdowns, and travel bans—the global median was 42.54, while Chile scored 58.1. This placed Chile 36th among 184 countries and 4th in South America, following Peru (60.1), Suriname (60.5), and Venezuela (80.5).

Within the period analyzed (2016-2022), an important cycle of protests known as the "social outburst" took place between October 2019 and March 2020 [@somma_no_2021]. However, the protest drastically decreased as a result of the lockdown [@joignant_informe_2020]. In terms of public views on economic inequality, it has been documented that critical views toward inequality increased in 2019 and deepened under the pandemic [@coes_radiograficambio_2023], which we argue can be in part explained by the increased risk exposure and vulnerability in the context of the COVID-19 pandemic. 
Hereby, a relevant aspect is the transformation of the labor market and its impact on income distribution [@ministerio_de_desarrollo_social_y_familia_resultados_2022;@instituto_nacional_de_estadisticas_resultados_2022,]. In 2015, the poverty rate was 11.2%, decreasing to 8.5% by 2017, with the 10/10 ratio at 31.6. In 2020, during the COVID-19 pandemic, poverty increased to 10.7%, and the 10/10 ratio surged to 301.7, while unemployment reached 10.5%, the highest level in a decade. By 2022, poverty declined to 6.5%, the 10/10 ratio improved to 61.2, and unemployment decreased to 7.9%, rising slightly to 8.7% in 2023.

Although the pandemic might have triggered such concerns mostly regarding health, we believe that other policy areas, such as education, could have been affected in this period. Consistently, our first hypothesis is:

$H_1$: The justification of educational inequalities decreases after the pandemic

Our second argument is based on the evidence that relates rational interests with justifications and preferences [@sonhing_failure_2019]. Research about the justification of educational inequality usually focuses on differences between status groups, based on the rational premise that those who are worst-off would tend to show lower levels of justification [@stoilovaFairnessEducationalOpportunities2023]. For instance, @valantPoliticsAchievementGaps2016 find that Americans are more concerned about—and more supportive of proposals to close—wealth-based achievement gaps than Black-White or Hispanic-White gaps. However, most of the research in this area goes beyond status differences, focusing on how subjective factors such as perceptions and beliefs are related to different levels of justification. In line with this evidence, we expect that:

$H_2$: Those with lower status characteristics show lower justification of educational inequalities

Thirdly, we address inequality perception. As pointed out by @janmaatSubjectiveInequalityReview2013, justifications are usually confused under the labels of “beliefs” and “attitudes,” and their distinction from perceptions is relevant as it helps state that justifying a certain level of inequality requires taking into account which level of (perceived) inequality is first and foremost being recognized [@sonhing_merit_2011;@castillo_multidimensional_2023]. For instance, a recent study by @dayPerceivedIdealInequality2023 on the perception and justification of inequality in university endowments in the U.S. points out that individuals tend to underestimate the magnitude of existing inequality while simultaneously desiring greater equality. Furthermore, through an experiment, they demonstrate that information about inequality in university endowments (i.e., manipulating inequality perception) increases the perception of injustice in education. Results of this study are in line with the consistent evidence on biases related to the underestimation of economic inequality [@castillo_perception_2022;@gimpelson_misperceiving_2018]. As a larger inequality in this domain threatens a basic equality ideal, the corresponding hypothesis in this case is:

$H_3$: A larger inequality perception is negatively associated with the justification of educational inequalities.

Our fourth hypothesis refers to the central concept of meritocracy. As meritocracy justifies inequalities based on performance, we expect that:

$H_4$: Higher meritocratic perceptions are positively related to the justification of educational inequalities.

Finally, and in line with hypotheses 1 and 4, we explore interactions between meritocracy and inequality justification over time. As COVID-19 pandemic has affected the structural opportunities to attend schools, universities, or workplaces, the ideal of meritocracy might have been challenged in the time frame analyzed. As a consequence, the justification of inequality could show a decrease over time:

$H_5$: The positive association between meritocracy and inequality justification mitigates along time