### 2025-01-17 03-methods.qmd
methods & descriptives

#  Data, variables & methods

## Data

The main data source is the Chilean Longitudinal Social Survey ELSOC (2016–2023). ELSOC has been designed as a yearly panel study to evaluate how individuals think, feel, and behave regarding a set of social issues concerning conflict and social cohesion in Chile. The sampling design is probabilistic, stratified, clustered, and multistage. It provides adequate coverage of the country's largest cities (Metropolitan Area of Santiago, Valparaíso, and Concepción) and smaller cities. The sample covers participants aged between 18 and 75 years (in wave 1), representing  77% of the total national population and 93% of the urban areas [@elsoc_estudio_2022].

The survey has been conducted yearly since 2016, except for the year 2020, when it was suspended due to the pandemic. The waves 2016, 2017, 2018, 2019, 2022, and 2023 were administered using computer-assisted personal interviewing (CAPI). However, a reduced version was conducted using computer-assisted telephone interviews (CATI) in 2021. In addition, wave 3 included a refreshment sample to counter survey attrition. As a result, the total sample of wave 3 included 3,748 cases, of which 2,229 were part of the original sample and 1,519 were from the refreshment sample. The data from the refreshment sample is not included in the analytical sample for this study as we analyze longer response trends. Regarding the original sample, the response rate was 62.4% in wave 1, achieving N = 2,927 participants. In broader terms, the accumulated attrition between wave 1 and wave 7 is 40%, achieving a final sample of N = 1,741. The analysis used longitudinal weights to avoid bias due to the sampling design, allowing us to control for biases arising from systematic patterns of non-participation in the survey after the first wave. For a more detailed analysis of responses, attrition, and the construction of longitudinal weights, visit https://coes.cl/encuesta-panel/. The dataset is publicly available at: https://dataverse.harvard.edu/dataverse/elsoc.

## Variables

The dependent variable of this study is the justification of educational inequality. This construct is measured using the following statement: "It is just that high-income people have a better education for their children than people with lower incomes" ("Es justo que las personas de altos ingresos tengan una mejor educación para sus hijos que las personas con ingresos más bajos" in Spanish). Here, the respondents declared their preferences on a Likert scale from (1) "strongly disagree" to (5) “strongly agree” (5). The main independent variables refer to meritocratic perception, which is operationalized by two items, one related to effort ("In Chile, people are rewarded for their efforts"), and the other related to talent ("In Chile, people are rewarded for their intelligence and skills"). In addition, we included an indicator for economic inequality perception ("In Chile, the income differences are too large"). @fig-frecuencies shows the average distribution throughout the years for these variables. We observe that there is a large group (88.2%) of respondents who disagree or strongly disagree that people with higher incomes can access better education. Likewise, a large group of respondents (91.7%) agree or strongly agree that income differences in Chile are too large. On the other hand, there is an important level of disagreement (disagree + strongly disagree) about the perception of meritocracy, this is, whether people are rewarded for their talent (47.6%) and for their efforts (57%).

```{r echo=FALSE, fig.align='center'}
#| label: fig-frecuencies
#| fig-cap: "Frequency of responses for justification of educational inequality, perception of inequality, and perception of meritocracy for all the years analyzed"

 knitr::include_graphics("output/graphs/merit_plot.png")
```

A third group of variables is included in the analysis to assess the effect of elements related to the pandemic context. These variables were only included in the 2021 wave, so each individual's responses were replicated for the remaining years. The first variable addresses the belief regarding whether, in the context of the pandemic, it is more important to give priority to economic activity than to the health of the population, measured on a Likert scale from (1) "strongly disagree" to (2) "strongly agree". The second variable is a dummy (No/Yes) that captures whether individuals received state benefits during the pandemic (e.g., Food Box, Emergency Family Income, Contribution for the Middle Class, or others).

For testing the hypothesis regarding status we considered educational level and household income. Also, we included subjective social status as it has been argued that perceived social status is a relevant predictor of attitudes toward economic inequality [@schneider_poverty_2015;@castillo_meritocracia_2019]. In addition, such subjective measures can complement objective measures in predicting life chances [@oesch_subjective_2023], which are arguably connected to how individuals experience economic inequality. In this regard, it has been argued that how people form their views on their standing in society results from experiences with direct socioeconomic circumstances and social comparison processes with reference groups [@Condon2020]. Finally, political identification on a left-right scale, age groups, and gender were included as control variables. @tbl-descriptives shows the whole set of independent variables, the corresponding items,  response categories, and frequencies.

```{r echo=FALSE, results='asis'}
#| label: tbl-descriptives
#| tbl-cap: "Independent variables"

pacman::p_load(kableExtra)
kable(data.frame(image='![](output/tables/desc02.png)'),    # generar tabla
      format="pipe",
             col.names = "",
             row.names = F)
```

## Methods

Given the hierarchical structure of the data (observations nested in survey waves), we estimated a series of longitudinal multilevel models [@singer_applied_2009]. Such an approach is suited to account for the shared variance among units (in this case individuals), adjusting the estimation of the standard errors. The linear multilevel models are estimated using the R library "lme4" [@singer_applied_2009, p.4].