### 2025-01-19 04-analisis.qmd
analysis & discussion

# Results

Beginning with a descriptive analysis of changes in the justification of educational inequality between 2016 and 2023, @fig-alluvial illustrates yearly frequencies and transitions between years. Each year represents stacked porcentual frequencies, whereas the flows in between reflect the within-subject change of opinions from one year to the next [@rosvall_mapping_2010]. For instance, of the 32.9% who strongly disagreed with inequality justification in 2016, about half of them kept responding the same in 2017, whereas the other half shifted their opinion to other response categories. In general, the large majority - between 80 and 90% - disagrees with the justification of educational inequality throughout the years. Despite this overall tendency, we also observe that the disagreement with inequality justification (disagree + strongly disagree) tends to go down in the last waves. This change is mostly a result of the increase in the "agree" category, which doubles compared to previous years (from 5.4% in 2016 to 11% in 2023).

```{r echo=FALSE, fig.align='center'}
#| label: fig-alluvial
#| fig-cap: "Change in the justification of educational inequality over time (2016-2023)"
#| 
 knitr::include_graphics("output/graphs/alluvial_dep.png")
```

@fig-means shows the average changes in the main variables considered for this study. Here we observe that the justification of educational inequality has the lowest average throughout the years when compared with the other (independent) variables, whereas the highest average is consistently represented by inequality perception. Interestingly, in the last waves of the study (2022-2023) the justification of inequality increases whereas the perception of inequality decreases. As the merit variables are concerned, they show a very similar pattern in terms of averages and changes over the years, being the perception of meritocracy related to effort always lower than the one associated with talent.

```{r echo=FALSE, fig.align='center'}
#| label: fig-means
#| fig-cap: "Change in the mean of justification of educational inequality, perception of inequality, and perception of meritocracy over the years"

 knitr::include_graphics("output/graphs/years_plot.png")
```

@fig-correlation presents a correlation matrix of the main variables analyzed, using data from the first wave. In this matrix, the correlations vary between low and moderate values. The correlations between justification of educational inequality with the two COVID variables are positive, but low. In addition, justification of education inequality depicts a moderate and negative association with the perception of inequality (_r_=-0.34, p<.01) and a moderate and positive association with both meritocracy perception variables (_r_=0.29, p<.01; _r_=0.24, p<.01). Regarding the perception of inequality, it presents a moderate and negative association with both meritocracy perception variables (_r_=-0.18, p<.01; _r_=-0.08, p<.01). Finally, the two meritocracy variables present a high and positive association with each other (_r_=0.76, p<.01).

```{r echo=FALSE, fig.align='center'}
#| label: fig-correlation
#| fig-cap: "Correlation matrix of the main variables for the first wave (2016)"

 knitr::include_graphics("output/graphs/corr.png")
```

_Multilevel models_

@tbl-multilevel shows the multilevel estimation results for the justification of educational inequality. Model 1 includes the survey waves to estimate intertemporal changes in the dependent variable. Taking 2016 as a reference point, we can observe a staggered decrease in 2017 ($\beta$=-0.167, p<.001), 2018 ($\beta$=-0.059, p<.05), and 2019 ($\beta$=-0.03, p>.05). Nevertheless, in the last waves of 2022 and 2023 there is a clear increase in level of justification of educational inequality ($\beta$=0.204, p<.001 and $\beta$=0.233, p<.001), suggesting a non-linear change in this variable. Attempting to model this path of change over time, Model 2 incorporates time (survey waves) as a continuous variable as well as its quadratic term, representing the nonlinear association initially observed in Model 1. On the one hand, the survey wave depicts a negative association, expressing an average decrease in inequality justification over time. On the other hand, the quadratic wave term is positive, indicating the reversion of this path in the last measurement points.

Models 3 and 4 adds the two variables about the pandemic context, where we can see that those who received state benefits during this period justify less inequality in education compared to those who did not receive them. Although the effect size is small, it remains significant when controlling for the rest of the variables in model 8 ($\beta$=-0.093, p<.01). On the other hand, those who believe that in this context the economy is more important than the health of the population justify more inequality in education($\beta$=0.038, p<.05, Model 8).

Models 5 and 6 introduce the meritocratic variables: talent (if intelligence and abilities are rewarded in society) and effort (if efforts are rewarded in society). In line with our hypotheses, the perception that talent is rewarded has a positive influence on the justification of educational inequality in Model 5 ($\beta$=0.055, p<.001). However, when controlling for the perception that effort is rewarded, this effect is no longer significant. In this sense, Model 6 shows that the perception that effort is rewarded in society is not only positively associated with the justification of educational inequality ($\beta$=0.083, p<.001), but it has a larger weight than the perception of talent. Regarding inequality perception (added in Model 7), it depicts a negative association with the justification of educational inequality, remaining stable when controlling for the rest of the variables.

Model 8 adds the socioeconomic variables to test hypothesis 7. Although educational level and household per capita income are not significant in this model, controlling for all variables in the next models, household per capita income depicts a positive effect ($\beta$=0.001, p<.05). The inclusion of other controls as being right-wing (compared to being left-wing) shows a positive association with justification of educational inequality ($\beta$=0.27, p<.001), whereas being a women (compared to men) had a negative  effect ($\beta$=-0.08, p<.001). Subjective social status and age had no significant effect when controlling for all the variables analyzed. The full model specification can be seen in the appendix (see @tbl-multilevel-full).

```{r echo=FALSE, results='asis'}
#| label: tbl-multilevel
#| tbl-cap: "Multilevel longitudinal models for the justification of inequality in education"

pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               summarytools,
               lme4,
               texreg,
               ggeffects,
               ggplot2
               )
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))

# generate analytical sample
df_study1 <- 
  df_study_long_t7_comunas %>%
  select(idencuesta,ola,comuna, ponderador_long_total,just_educ, des_perc, merit_effort, merit_talent, education=educ,quintil1:sexo, covid_economia, covid_beneficios, ing_pc) %>% 
  na.omit() %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2,
         sexo=as.factor(sexo),
         ing_pc_rec = (ing_pc/100000)
         )

df_study1 <- df_study1 %>%
  group_by(idencuesta) %>%             # Agrupar por el identificador del participante
  mutate(n_participaciones = n()) %>%  # Contar el número de filas (participaciones) por participante
  ungroup()
#sjmisc::frq(df_study1$n_participaciones)
df_study1 <- df_study1 %>% filter(n_participaciones>1)

#, percentile_wave, bot_mid_top_wave, inc_pcmed, ratioincome, incomegrp
df_study1$covid_beneficios<- car::recode(df_study1$covid_beneficios, c("1=2; 2=1"))
df_study1$covid_beneficios<- factor(df_study1$covid_beneficios)

# Hipotesis
h0 <- "ola"
h1 <- "ola_num+ola_2"
h2 <- "ola_num+ola_2+covid_beneficios"
h3 <- "ola_num+ola_2+covid_beneficios+covid_economia"
h4 <- "ola_num+ola_2+covid_beneficios+covid_economia+merit_talent"
h5 <- "ola_num+ola_2+covid_beneficios+covid_economia+merit_talent+merit_effort"
h6 <- "ola_num+ola_2+covid_beneficios+covid_economia+merit_talent+merit_effort+des_perc"
h7 <- "ola_num+ola_2+covid_beneficios+covid_economia+merit_talent+merit_effort+des_perc+education+ing_pc_rec+ess+edad+sexo+pos_id"

h8 <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort"

# A. Education distributive justice
educ.null <- lmer(formula(paste0("just_educ~","1 + (1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ0 <- lmer(formula(paste0("just_educ~",h0,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ1 <- lmer(formula(paste0("just_educ~",h1,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ2 <- lmer(formula(paste0("just_educ~",h2,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ3 <- lmer(formula(paste0("just_educ~",h3,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ4 <- lmer(formula(paste0("just_educ~",h4,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ5 <- lmer(formula(paste0("just_educ~",h5,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ6 <- lmer(formula(paste0("just_educ~",h6,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ7 <- lmer(formula(paste0("just_educ~",h7,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ8 <- lmer(formula(paste0("just_educ~",h8,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)

# Table 1: individual

omit <- "(edad)|(ess)|(sexo)|(pos_id)"
texreg::knitreg(list(educ0, educ1, educ2, educ3, educ4, educ5, educ6, educ7)
                     ,
                digits = 3,
                custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05.
Note: Model 8 is controlled by age, gender, political position, and subjective social status.", 
                omit.coef = omit,
                custom.coef.names = c("Intercept",
                                      "Wave 2017",
                                      "Wave 2018",
                                      "Wave 2019",
                                      "Wave 2022",
                                      "Wave 2023",
                                      "Wave",
                                      "Wave^2",
                                      "Covid: receive state benefits",
                                      "Covid: economy more important than health",
                                      "Merit: Talent",
                                      "Merit: Effort",
                                      "Inequality perception",
                                      "Universitary (Ref.= Less than universitary)",
                                      "Household income"
                                      # "High School",
                                      # "Technical",
                                      # "Universitary",
                                      # "Quintile 2",
                                      # "Quintile 3",
                                      # "Quintile 4",
                                      # "Quintile 5",
                                      # "Quintile (No information)",
#                                      "Subjective Social Status",
                                      # "Age 30-49",
                                      # "Age 50-64",
                                      # "Age 65 or more",
                                      # "Gender (Ref. Male)",
                                      # "Center",
                                      # "Right",
                                      # "Does not identify",
                                      ),
                groups = list("Wave (Ref.= Wave 2016)" = 2:6
                          #    "Education (Ref.= Primary)" = 9:11,
                  #"Household Income (Ref.= Quintile 1)" = 12:16
                  #"Age (Ref.= 18-29)" = 14:16,
                  #"Pol. pos (Ref.= Left)" = 18:20
                  ),
                include.loglik = FALSE,
                include.aic = FALSE,
                custom.gof.rows = list("Controls"= c(rep("No", 7), "Yes")),
                custom.gof.names = c("BIC",
                                     "Num. obs.",
                                     "Num. groups: Individuals",
                                     "Var: Individuals (Intercept)",
                                     "Var: Residual"),
               center = TRUE,
               caption.above = TRUE,
               caption = "", 
               scalebox = 0.765,
               float.pos = "!ht", 
               booktabs = FALSE, 
               use.packages = FALSE, 
               include.rs = FALSE, 
               include.rmse = FALSE,
               include.ci = FALSE)
```

In this last part of the analysis we test hypotheses about the effect of COVID-19 on changes in the relationship between meritocracy and justification of educational inequality. In hypothesis 5, we proposed that the association between the perception of meritocracy and justification of inequality is mitigated in times of crisis, as meritocratic ideals might have been weakened due to critical situations associated with the COVID-19 pandemic. We test this hypothesis through interaction effects, shown in @tbl-interact. The first model is shown as the baseline model. It is the same as model 8 in @tbl-multilevel, but only shows the variables involved in the interaction for the sake of space (all other variables are included). Model 9 adds the interaction between the perception of talent-based meritocracy and whether state benefits were received during the pandemic, whereas Model 10 considers the perception of effort-related meritocracy for the interaction. Both variables show a negative significant interaction with receiving state benefits ($\beta_{talent}$=-0.07, p<.01; $\beta_{effort}$=-0.05, p<.05), meaning that for those who received state benefits the association between the the justification of inequality and the perception of effort-based meritocracy lower by 0.05 points, and by 0.07 points regarding the perception of talent-based meritocracy. This last interaction is depicted in @fig-interact-talent, showing that the positive association between the perception of talent-based meritocracy and the justification of educational inequality is smaller for those who received state benefits during the pandemic.

```{r interact, echo=FALSE, results='asis', warning=FALSE}
#| label: tbl-interact
#| tbl-cap: "Interaction effects for the justification of economic inequality"

# df_study1 <- 
#   df_study_long_t7_comunas %>%
#   select(idencuesta,ola,comuna, ponderador_long_total,just_educ, des_perc, merit_effort, merit_talent,education=educ,quintil1:sexo, covid_economia, covid_beneficios) %>% 
#   na.omit() %>% 
#   mutate(ola = as.numeric(ola),
#          ola_2=as.numeric(ola)^2,
#          sexo=as.factor(sexo)
#  #        edad=as.numeric(edad)
#          #pos_id=as.numeric(pos_id)
#          )

# df_study1$ola <- car::recode(df_study1$ola, c("1=1; 2=1;3=1;4=1; 6=2;7=2"))
# df_study1$ola <- as.factor(df_study1$ola)
#df_study1$covid_beneficios<- car::recode(df_study1$covid_beneficios, c("1=2; 2=1"))
# Hipotesis

h6 <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort"
h1b <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+merit_effort*covid_economia"
h3b <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+merit_talent*covid_economia"
h4b <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+des_perc*covid_economia"
h5b <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+merit_effort*covid_beneficios"
h6b <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+merit_talent*covid_beneficios"
h7b <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+des_perc*covid_beneficios"
# A. justification educational inequality
educ0 <- lmer(formula(paste0("just_educ~",h6,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ1 <- lmer(formula(paste0("just_educ~",h1b,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ3 <- lmer(formula(paste0("just_educ~",h3b,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ4 <- lmer(formula(paste0("just_educ~",h4b,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ5 <- lmer(formula(paste0("just_educ~",h5b,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ6 <- lmer(formula(paste0("just_educ~",h6b,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ7 <- lmer(formula(paste0("just_educ~",h7b,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)

# table 3: interacciones
omit <- "(education)|(ing_pc_rec)|(ess)|(edad)|(sexo)|(pos_id)|(covid_economia)|(des_perc)"
texreg::knitreg(list(educ0,educ6, educ5),
                custom.note = "\\item *** p < 0.001; ** p < 0.01; * p < 0.05. Note: All the models are controlled by educational level, income quintile, subjective social status, age, gender, political position, and inequality perception",
                omit.coef = omit,
                include.loglik = FALSE,
                include.aic = FALSE,
                custom.model.names = c("Model 8", "Model 9", "Model 10"),
                custom.coef.names = c("Intercept",
                                      "Wave",
                                      "Wave^2",
                                      "Covid: receive state benefits",
                                      "Merit: Talent",
                                      "Merit: Effort",
                                      "Receive state benefits * Merit: Talent",
                                      "Receive state benefits * Merit: Effort"
                                      ),
                custom.gof.names = c("BIC",
                                     "Num. obs.",
                                     "Num. groups: Individuals",
                                     "Var: Individuals (Intercept)",
                                     "Var: Residual"),  
               center = TRUE,
               caption.above = TRUE,
               caption = "",
               scalebox = 0.765,
               float.pos = "!ht", 
               booktabs = TRUE, 
               use.packages = FALSE, 
               include.rs = FALSE, 
               include.rmse = FALSE,
               include.ci = FALSE)
```

```{r echo=FALSE, fig.align='center'}
#| label: fig-interact-talent
#| fig-cap: "Interaction effect of merit based on talent and receive state benefits on justification of educational inequality"

ggeffects::ggpredict(educ6, terms = c("merit_talent", "covid_beneficios")) %>%
  ggplot(aes(x=x, y=predicted, shape = group, color = group)) +
  geom_line(aes(group=group,linetype = group),position = position_dodge(.1)) + 
  geom_point(size = 2.5,position = position_dodge(.1))+
  scale_y_continuous(limits=c(1,5),  breaks=c(1,2,3,4,5))+
  scale_x_continuous(name = "Merit: talent", breaks=c(1,2,3,4,5), labels = c("Strongly
disagree", "Disagree", "Neither agree
nor disagree", "Agree", "Strongly
agree")) + 
  scale_shape_discrete(name = "Receive state
benefits",
                       limits = c("1", "2"),
                       labels = c("No", "Yes")) +
  scale_color_manual(name = "Receive state
benefits",
                     limits = c("1", "2"),
                     labels = c("No", "Yes"),
                     values = c("black", "black")) +
  scale_linetype_manual(name = "Receive state
benefits",
                        limits = c("1", "2"),
                        labels = c("No", "Yes"),
                        values = c("solid", "dashed"))+
  theme_bw() +
  labs(title="", y = "Predics values of justification of educational inequality") + 
  theme(plot.title = element_text(size = 11),
        axis.text=element_text(size=11))
```