### 2025-01-19 05-discussion.qmd

# Discussion and conclusions

Our research sought to examine the justification of inequality in education from a longitudinal perspective amid the health crisis generated by the COVID-19 pandemic. In general, our results showed mixed evidence about our hypotheses. We argued that in times of crisis, greater exposure to risk due to the health context would result in a lower justification of inequality by citizens. In this regard, longitudinal models showed that the justification of inequality in education is far from being linear, and whereas a decreasing pattern was found between 2017 and 2019, there is a striking increase in the post-pandemic measurement points. Besides, our findings revealed that those who received benefits justified educational inequality less than those who did not. Still, given our data limitations it is not possible to attribute such changes only to the pandemic and the related economic and social crises. As during this very same period in Chile there were two (failed) constitutional processes, disentangling different forces driving public opinion becomes cumbersome and deserves further qualitative and quantitative studies. Along this line, some recent studies have argued that the election of a far left-wing constitutional assembly in the first process generated a backslash effect, mainly due to several scandals that led to the delegitimation of this assembly. This could have driven preferences in a conservative direction, which was reflected in the election of a right-wing second assembly after the failed first constitutional process [@palanza_chiles_2023;@sazo_chile_2023].

The main focus of this study was on the relationship between meritocracy and justification of inequality, arguing that those who perceive that the society in which they live complies with meritocratic principles, would be more inclined to justify inequality in education. At first look, the results are consistent with the hypothesis; however, some nuances are worth attending as the effort dimension prevails over the talent dimension. In other words, the perception of meritocracy in terms of rewarded effort is more relevant in justifying access to education than the perception of talent. This could be explained as that talent could be more associated with luck in terms of random assignment, and in this terms would not be enough reason (as it is effort) to justify educational inequality. Furthermore, we observed that the connection between perceived meritocracy and the justification of educational inequality weakened for those who received state benefits, which aligns with our expectations given the social context in the post-COVID era. One possible interpretation is that, during times of increased risk, people may become more skeptical of narratives emphasizing individual effort over collective efforts. This shift in perspective might also foster stronger solidarity to protect individual and societal wellbeing. However, further studies are necessary to explore this phenomenon using more specific conceptual and methodological approaches.

Regarding the status position, our central hypothesis was that individuals in more advantaged situations and with greater resources would tend to defend their interests and justify greater education inequality. In this regard, our results show a positive relationship between income and the justification of inequality, which would align with our argument that rational interests of the better-off that would lead to a larger inequality justification. Besides status, we also considered inequality perception, as the literature on attitudes toward inequality has argued that a greater justification of inequality is associated with its perceived magnitude. Consistent with this hypothesis, our results showed that, throughout the years, a larger perceived economic inequality motivates a lower justification of inequality in education.

Among the limitations of our study, we can mention at least three. First, the operationalization of justification of educational inequality is rather limited. We counted only on a single item for this in this study, and further work is needed in order to assess the validity and reliability of measuring such construct. Second, as it has been established by recent research [@castillo_multidimensional_2023], the perception of meritocracy can be understood through individual attributions regarding effort and talent, but also taking into account the role of structural aspects such as family status and social capital. Due to data limitations such dimensions were not included here, therefore providing a partial perspective on the areas of perceived meritocracy. Finally, there are limitations involving the longitudinal dimension of the data used. Given constraints for data collection during the pandemic, the ELSOC study had to shorten the questionnaire in 2021 and unfortunately, the educational justice indicator was excluded from that survey. Consequently, the results of temporal change have to be considered carefully, and further data waves could give us more information on this matter.

Regarding future studies, the capabilities of the ELSOC longitudinal database are not limited to micro-level estimates. In this regard, thanks to the sampling strategy of the survey, it would be possible to make contextual estimates at the municipality level. For instance, statistical models could include contextual information from administrative data sources, allowing testing for hypotheses that include contextual socioeconomic inequality as well as its dynamics over time.  Furthermore, this dataset could be used for comparing inequality justification in policy areas besides education, such as health and pensions. Future studies may shed light on how meritocratic perceptions and beliefs affect differently such areas and their changes over time, which would give relevant hints from social sciences for the discussion about societal and cultural changes as well as their impacts on solidarity and social cohesion.